Molecular Gas Dynamics of the Strongly-lensed Quasar Host Galaxy RXS J1131$-$1231

### 2016-08-28 WRITEUP/ms.tex
added DR's comments, major change: gas fraction, sSFR, intro and conclusion, and removed 'nuclear SB'

\citestyle{aa}
\shorttitle{Molecular Gas in the Lensed Wet-Merger RXJ1131$-$1231 at $z$\,=\,0.65}
\shortauthors{Leung, Riechers \& Pavesi}

@@

\title{Molecular Gas Dynamics of the strongly-lensed wet-merger RXS J1131$-$1231 at $z$=0.654}
\author{T. K. Daisy Leung, Dominik A. Riechers, and Riccardo Pavesi}
\affil{Department of Astronomy, Space Sciences Building, Cornell University,
Ithaca, NY 14853, USA; \myemail}

@@

\begin{abstract}
We report observations of the \bco and \cco line emission towards
the quadruply-imaged quasar
RXS\,J1131$-$1231 at $z$\,=\,0.654 obtained using the
\pdbi (PdBI) and \carma (CARMA), making this the first spatially resolved
CO study at intermediate redshift.
Our lens modeling shows that the highly asymmetric double-horned line profile
observed in the \bco spectrum is mainly a result of differential lensing, where
the magnification factor varies from $\sim$3 to $\sim$8 across
different kinematic components.
The intrinsically symmetric line profile with a clear velocity
gradient in the source-plane suggests that
the quasar host galaxy is a rotating disk.
In addition, we find a secondary CO-emitting source near RXJ1131 whose location is
consistent with the optically-faint companion reported in previous studies.
The lensing-corrected gas mass of
$M_{\rm gas}$\ssim1.4\E{10}\,\Msun inferred for RXJ1131
and \ssim2\E{9}\,\Msun for its companion correspond
to a gas mass ratio of $\sim$7:1.
The extended CO gas disk in RXJ1131 ($R$$\sim$6\,kpc) yields a dynamical mass
of $M_{\rm dyn}$\ssim8\E{10}\,\Msun, and thus a gas mass fraction of $\sim$19\%,
which is consistent with the trend of decreasing molecular gas fraction since $z$\ssim1.
The lensing-corrected $M^{*}$\ssim3\E{10}\,\Msun,
$M_{\rm dust}$\ssim4\E{8}\,\Msun, $L_{\rm IR}$\ssim1.5\E{12}\,\Lsun, and SFR$_{\rm FIR}$\eq120\pmm63 \sfrU of RXJ1131
together with the source-plane morphology of the gas and the stellar populations
suggest that ongoing \SBs are embedded within the molecular gas reservoir,
and that the nuclear \SBs are likely triggered by
interactions with the companion.
The $M_{\rm BH}$$/$$M_{\rm bulge}$ of $\gtrsim$0.27\% derived for RXJ1131 lies above the
local relation but is consistent with other radio-loud galaxies at this epoch, suggesting
that the co-evolution between RXJ1131 and its host galaxy is different than those of local AGN host galaxies.
\end{abstract}

@@

%--------------------------------------------------------------------------
%                                Introduction
%--------------------------------------------------------------------------
\section{Introduction}
Studies of galaxy evolution have been focusing on the interplay between
\SF and active galactic nucleus (AGN) activity across cosmic epochs.
It is not well-understood when and how stellar populations of present-day galaxies were assembled,
but it is clear that both the co-moving \SF rate and black hole accretion rate densities
increased substantially since $z$\,$>\,3$ and reached their climax at $z$\ssim$2$, followed by
a rapid decline toward $z$\ssim$0$ \citep[\eg][]{Hopkins06a, Madau14a}.
% Studying the cold gas at the redshift range 0.2 $<$\,$z$\,$<$\,1 of paramount importance to ravel the cause for this decline.
A leading explanation for this decline is related to the evolution of molecular gas content and
\SF efficiency \citep[]{Erb06a, CW13, Walter14a},
but confirmation from direct molecular gas measurements at intermediate redshift
$(0.2\,<\,z\,<\,1)$ is still limited both in number and in the variety of galaxy populations probed ---
only unresolved CO observations have been carried out for $\sim$$30$ ULIRGs up to date \citep{Combes11a, Combes13a}. %18+15

% In the current paradigm, it is posited that every nearby, massive galaxy host a supermassive black hole (SMBH), which
% grows by accretion of the cold molecular gas from its host galaxy.
Meanwhile, empirical scaling relations such as the
$M_{\rm BH}$$-$\mbulge relation \citep[e.g.,][]{Magorrian98a, HR04a}
has been established locally, suggesting a co-eval growth between local SMBHs and their host galaxies.
Attempts to extend this relation out to higher redshifts, beyond the peak epoch
of \SF and AGN activity, have been made in recent years.
It has been reported that the $M_{\rm BH}$$/$$M_{\rm bulge}$ ratio of high-$z$ submillimeter galaxies (SMGs) falls below
the local relation \citep[\eg][]{Borys05a,Alexander08a} but that of high-$z$ quasar host galaxies is found
lying above
the relation \citep[\eg][]{McLure06a,Peng06a}.
Stellar mass of quasar host galaxies at $z$\,$>$\,4 constrained by their CO-based dynamical mass
is also too small compared to that based on the local relation \citep{Walter04a, Riechers08a,Coppin08a}.
While it has become clearer that high-$z$ galaxies deviate from the local relation, and among their galaxy populations,
such a picture is still unclear  at intermediate redshift.
This is owed to the difficulties in measuring stellar emission in the host galaxy under the presence of a bright quasar, or alternatively
due to the lack of spatially resolved CO observations to provide dynamical constraints on the stellar mass.
With gravitational lensing, emission of the background quasar and its host galaxy
is magnified and stretched out to different extents, and is being projected onto different locations in the image plane,
allowing spatially resolved measurements their emission in a
relatively uncontaminated fashion.

\defcitealias{Sluse03a}{S03}
The quasar RXS J113151.62$-$123158 (hereafter RXJ1131)
at $z_\textrm{s, QSO}$\eq$0.685$ \citep[hereafter S03]{Sluse03a} is a particularly well-suited source for
studying the evolution of molecular gas properties in quasar host galaxies and the
connection between SMBHs and their host galaxies at intermediate redshift
given its unique lensing configuration.
The stellar emission in the host galaxy of RXJ1131 is lensed into
an Einstein ring of radius $1$\farcs$83$
that is clearly separated from the quadruply imaged quasar emission \citep[hereafter C06]{Claeskens06a}.
We have thus carried out a detailed study of RXJ1131 as presented in this paper.
RXJ1131 is strongly lensed by an elliptical galaxy at $z_\textrm{L}$\eq$0.295$ \citepalias{Sluse03a}.
The moderate mass black hole $M_{\rm BH}$\eq$8$\E{7}\Msun \citep{Sluse12a} %based on H beta line
residing in RXJ1131 is rotating with a high spin parameter \citep[$a$\,$\sim$\,$0.9$;][]{Reis14a}, and
is radiating at an intrinsic bolometric luminosity of $L_{\rm bol, X}$\eq$1.3$\E{45}\, ergs\,s\pmOne \citep{Pooley07a}.

This paper is structured as follows.
In \Sec{obs}, we outline details of the observations and of our data reduction procedures.
In \Sec{HST}, we describe the astrometric corrections
applied to align an {\it HST} image with an VLA radio image.
In \Sec{results}, we report measurements of the \bco and \cco lines and broadband
photometry spanning the optical to radio wavelengths.
In \Sec{anal}, we present dynamical lens modeling of the \bco data and
SED modeling of the host galaxy of RXJ1131 to
derive physical parameters reflecting the nature of this system.
In \Sec{diss}, we discuss the merger stage of RXJ1131,
the velocity offset between the AGN lines and the CO lines,
and the $M_{\rm BH}$$-$\mdyn relation of RXJ1131.
Finally, we summarize the main results and present our conclusions in \Sec{sum}.
We use a concordance $\Lambda$CDM cosmology throughout this paper, with
parameters from the WMAP9 results:
$H_0$ = $69.32$ \kms Mpc\pmOne, $\Omega_{\rm M}$ = $0.29$, and
$\Omega_{\Lambda}$ = $0.71$ \citep{Hinshaw13a}.
% same as LR16a: i.e. use last column (WMAP+eCMB+BAO+H0) of Tab 4 in Hinshaw+13 because that's what astropy use for WMAP9 as cosmo
\defcitealias{Claeskens06a}{C06}

%--------------------------------------------------------------------------
%                          Observations details
%--------------------------------------------------------------------------
\section{Observations} \label{sec:obs}
\subsection{PdBI \bco}
Observations of the \bco rotational line
($\nu_{\rm rest}$\,=\,$230.538$ GHz) redshifted to $\nu_{\rm obs}$\ssim$139.0$\,GHz
were carried out using IRAM \pdbi (PdBI; Program ID: S14BX; PI: D.
Riechers). Two observing runs were carried out on 2014 December 06 and 2015
February 05 under good weather conditions in the C and D array configurations,
respectively. The 2 mm receivers were used to cover the redshifted \bco line
and the underlying continuum emission, employing a correlator setup that provides
an effective bandwidth of 3.6 GHz (dual polarization) and a spectral resolution of 10.0 MHz ($\sim$
$21.5$ \kms). This resulted in 3.75 hours of cumulative six antenna-equivalent on-source
 time after discarding unusable visibility data.
The nearby quasars B1127$-$145 and B1124$-$186 were observed every 22 minutes
for pointing, secondary amplitude, and phase calibration, and B1055$+$018 was
observed as the bandpass calibrator for both tracks.
MWC349 and 3C279 were observed as primary flux calibrators for the C and D
array observations, respectively, yielding calibration accuracy better than 15\%.

@@

\subsection{CARMA \cco}
Observations of the \cco rotational line in RXJ1131
($\nu_{\rm rest}$\,=\,345.7959899\,GHz) redshifted to $\nu_{\rm obs}$\,=\,208.6\,GHz
were carried out with the \carma (CARMA;
Program ID: cf0098; PI: D. Riechers)
in the D array configuration on 2014 February 02 under poor 1.5\,mm
weather conditions and on 2014 February 17 under good 1.5\,mm
weather conditions. The correlator setup provides a bandwidth of 3.75 GHz in
each sideband and a spectral resolution of 12.5 MHz ($\sim$17.9 \kms). The
line was placed in the lower sideband with the local oscillator tuned to $\nu_{\rm LO}\sim$216 GHz. The radio quasars J1127$-$189 (first track) and 3C273
(second track) were observed
every 15 minutes for pointing, amplitude, and phase calibration. Mars was
observed as the primary absolute flux calibrator and 3C279 was observed as
the bandpass calibrator for both tracks. This results in a total on-source time of 2.94 hours after flagging poor
visibility data.

% poor phase note
% /Users/admin/Research/RXJ1131/CARMA/imagingD23/checkQuality.csh
% rms scatter over ~ 90 klambda
% max baseline of obs. is 145 m
%
% calflux source=1127-189 in=$MIRCAT/FluxSource.cat device=/xs
% Flux of: 1127-189  14FEB04.50 at 227.0 GHz:  0.65 Jy; rms: 0.10 Jy
% Extrapolate to 215.673 GHz with spectral index = -0.986 --> 0.684 Jy
% MIRIAD from bootstrap: Median Flux @ 215.673 GHz:     0.710 --> 3%
%
% amp de-correlation: % loss scales = 1- exp(-(50º*pi/180)^2 / 2) = ~32% loss
Given that the phase calibrator used for the first track was faint and was
observed under poor weather conditions and that the phase calibrator used for
the second track was far from our target source, the phase calibration is
subpar, with an rms scatter $\sim$50$\degr$ over a baseline length of $\sim$135\,m.
We thus conservatively estimate
a calibration accuracy of $\sim$35\% based on the flux scale uncertainties,
the gain variations over time, and the phase scatter on the calibrated data. We
therefore treat its line intensity with caution and ensure that our physical interpretation
of this system and the conclusion of this paper do not rely on this quantity.

@@

\subsection{VLA (Archival)} %DONE
Our analysis also uses archival data of the 5\,GHz
radio continuum obtained with the
Very Large Array (VLA; Program ID: AW741; PI: Wucknitz).
Observations were carried out on 2008 December 29 under excellent weather
conditions in the A array configuration for a total of $\sim$7 hours on-source time. The C-band receivers were used with a continuum mode setup,
providing a bandwidth of 50 MHz for the two IF bands with full polarization.
The nearby radio quasar J1130$-$149 was observed every 10 minutes for
pointing, amplitude, and phase calibration, J1331$+$305 was observed as the
primary flux calibrator, and J0319$+$415 was observed as the bandpass
calibrator, yielding $\sim$10\% calibration accuracy.
We used \ncode{aips} to calibrate the visibility data.
The calibrated visibility data were imaged and deconvolved using
the CLEAN algorithm with robust\,=\,0, which
was chosen to obtain a higher resolution image given high SNR.
This yields a synthesized clean
beam size of 0$\farcs$49 $\times$ 0\farcs35 (PA\,=\,0.18$\degr$) and a final
rms noise of $\sigma$ = 13 $\mu$Jy beam\pmOne.

\section{HST astrometry} \label{sec:HST}
We obtained an {\it HST} image taken with
the ACS/Wide Field Camera
using the F555W filter ($V$-band)
from the
Hubble Legacy Archive\footnote{Based on observations
made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble
Legacy Archive, which is a collaboration between the Space Telescope Science
Institute (STScI/NASA), the Space Telescope European Coordinating Facility
(ST-ECF/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).}.
The details of the observations can be found
in \citetalias{Claeskens06a}.
We adopt the VLA 5\,GHz map as the
reference coordinate frame to align the optical ($V$-band) image.
We shift the latter to the east by 0\farcs5963 in R.A. and $+$0\farcs8372 in
Dec., which is consistent with the typical astrometric precision (1$^{\prime\prime}-$2$^{\prime\prime}$) of
images from the Hubble Legacy Archive\footnote{http://hla.stsci.edu/hla\_faq.
html}. This astrometric correction is critical to avoid artificial spatial
offsets between different emitting regions and to carry out our lens modeling,
in which the absolute position of the foreground lensing galaxy is based on
its coordinates in the high-resolution optical image.
The VLA image is calibrated using a well-monitored phase
calibrator, with absolute positional accuracy of $\sim$2 mas.
For this reason, the absolute alignment between the VLA image and other
interferometric images reported in this paper are expected to have an astrometric
precision better than 0\farcs1, modulo uncertainties related to the SNR and phase
instability.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.45\textwidth]{../Figures/SpecCO21_twinx.eps}
\caption{Continuum-subtracted spectrum of \bco emission toward RXJ1131. The velocity scale
is with respect to $z$\,=\,0.6537, which is approximately the line center
considering the asymmetry resulting from differential lensing.
A detailed discussion of this effect is presented in
\Sec{differential}.
 \label{fig:CO21spec}}
\end{figure}

% moment 0 map, highest SNR: CASA chan=125~159 <=> GILDAS 126-160 <=> python 125-160
% - sigma = 0.305 Jy km/s/Beam
%   - note that theoretical sigma is lower = 1.5 * sqrt(160-126+1) * 21.5 ~ 0.2 Jy km/s /beam, due to higher noise in some channels with emission
We construct the zeroth order moment map, a red/blue channel map, and
the first and second moment maps in \Fig{CO21mom}
using the $uv$-continuum subtracted data cube over a velocity range of
$\Delta v$ $\sim$ 750\,\kms.
The higher-order moment maps are produced using
unbinned channel maps with 3$\sigma$ clipping.
The peak flux density is 8.12\pmm0.30 Jy\,\kms\,beam\pmOne
in the intensity map.

The deconvolved source size FWHM from fitting a two-dimensional Gaussian
is 5\farcs1\pmm0\farcs72 $\times$ 3\farcs72\pmm0\farcs66,
and thus, the emission is resolved over $\sim$2.2 beams.
While the lensed emission is not strictly distributed as a two-dimensional
Gaussian;
the fit recovers the line intensity enclosed by the emitting
region, therefore we take this as an estimate on the full extent of the lensed
emission. On the other hand, if we assume that the spatial distribution of
the lensed molecular gas emission is similar to that in the optical to \nir
wavelengths, the lensed emission would be more accurately described by an
annulus, enclosing the partially complete ``Einstein ring'' and
the lensed knots (see \Fig{CO21mom}).

\begin{figure*}[!htbp]
\hspace{0.5em}
\includegraphics[trim=0 10 15 0, clip, width=0.452\textwidth]{../Figures/{F555WCO21_mom0_single.invertedgray}.eps}
\includegraphics[trim=5 -18 0 0, clip, width=0.462\textwidth]{../Figures/F555W_REDBLUE.eps}
\\
\includegraphics[trim=25 16 0 10, clip, width=0.85\textwidth]{../Figures/CO_highOmom_CLIP5sigma}
\vspace{0.1em}
\caption{Top left: overlay of the velocity-integrated \bco emission on the archival {\it HST} $V$-band (F555W)
image.
Top right: same as top left, except the contours are color-coded to represent the red- and blueshifted emission.
The contours in both top panels start at 3$\sigma$ and increment at steps of
$\pm$3$\sigma$, where $\sigma$\,=\,0.3 mJy beam\pmOne for the top left panel,
and
$\sigma$ = 0.4 mJy beam\pmOne (red) and 0.5 mJy beam\pmOne (blue)
for the top right panel.
The crosses denote the
location of the foreground galaxy at $z$\,=\,0.295.
Contours for the first (bottom left) and second (bottom right) moment maps of the \bco line emission
are shown in steps of
50 \kms, and 100 \kms, respectively.
The synthesis beam size is 4\farcs4 $\times$ 2\farcs0, at PA = 13$\degr$.
\label{fig:CO21mom}}
\end{figure*}

>>>>

\citestyle{aa}
\shorttitle{Molecular Gas Dynamics in the Strongly-lensed Quasar Host Galaxy RXJ1131$-$1231}
\shortauthors{Leung, Riechers \& Pavesi}

@@

\title{Molecular Gas Dynamics of the Strongly-lensed Quasar Host Galaxy RXS J1131$-$1231}
\author{T. K. Daisy Leung, Dominik A. Riechers, and Riccardo Pavesi}
\affil{Department of Astronomy, Space Sciences Building, Cornell University,
Ithaca, NY 14853, USA; \myemail}

@@

\begin{abstract}
We report observations of \bco and \cco line emission towards
the quadruply-lensed quasar
RXS\,J1131$-$1231 (hereafter RXJ1131) at $z$\,=\,0.654 obtained using the
\pdbi (PdBI) and the \carma (CARMA).
Our lens modeling shows that the asymmetry of the double-horned \bco line profile
is mainly a result of differential lensing, where
the magnification factor varies from $\sim$3 to $\sim$8 across
different kinematic components.
The intrinsically symmetric line profile with a conspicuous source-plane velocity
gradient suggests that
the quasar host galaxy is an extended rotating disk, with a CO size of $R$$\sim$6\,kpc in radius and
a dynamical mass of $M_{\rm dyn}$\ssim8\E{10}\,\Msun.
In addition, we find a secondary CO-emitting source near RXJ1131 whose location is
consistent with the optically-faint companion reported in previous studies.
We decompose the lensing-corrected total \bco line flux into two components and 
find a lensing-corrected gas mass of $M_{\rm gas}$\ssim1.4\E{10}\,\Msun for RXJ1131
and \ssim2\E{9}\,\Msun for the companion, corresponding
to a gas mass ratio of $\sim$7:1.
The gas mass fraction of $\sim$19\% inferred for RXJ1131 is consistent with the trend of decreasing molecular gas fraction since $z$\ssim1 reported in recent studies.
The lensing-corrected stellar mass $M_{*}$\ssim3\E{10}\,\Msun,
dust mass $M_{\rm dust}$\ssim4\E{8}\,\Msun, IR luminosity $L_{\rm IR}$\ssim1.5\E{12}\,\Lsun, and 
star formation rate SFR$_{\rm FIR}$\eq120\pmm63 \sfrU of RXJ1131
together with its source-plane morphology of the gas and stellar populations
indicate ongoing star-formation activity that maybe as extended as the molecular gas reservoir.
We find a $M_{\rm BH}$$/$$M_{\rm bulge}$ ratio of 0.27\petm{11}{8}\% that is higher
than the local value
but consistent with other galaxies at this epoch. Collectively, our results suggest
that the bulk of the black hole mass of RXJ1131 is already in place 
while its stellar bulge component is still assembling, and that its major mode of \SF is likely merger-driven.
\end{abstract}

@@

%--------------------------------------------------------------------------
%                                Introduction
%--------------------------------------------------------------------------
\section{Introduction}
%Studying the interplay between \SF and active galactic nucleus (AGN) activity across cosmic epochs is crucial to 
%our understanding of the
%processes governing the formation and evolution of both the SMBHs and their host galaxies.
% Significant attention has been focused on studying the interplay between star formation and active galactic nucleus (AGN) activity across cosmic epochs in recent studies of galaxy evolution. 
% \citep{Wang10, Lamastra10, Hopkins+05, 06, DiMatteo+05a, Page11, Page12,Simpson12}
Many recent studies of galaxy evolution have been focused on investigating the interplay between star formation and active galactic nucleus (AGN) activity across cosmic epochs \citep[e.g.,][]{DiMatteo05a,Alexander05a,Hopkins06b,Coppin08a,Page12a,Simpson12a,Lamastra13a}. 
It is currently not well-understood when and how the supermassive black holes (SMBHs) and 
stellar populations of present-day massive galaxies were assembled,
but it is clear that the co-moving \SF rate and the black hole accretion rate densities both
increased substantially since $z$\,$>\,3$ and reached their climax at $z$\ssim$2$, followed by
a rapid decline toward $z$\ssim$0$ \citep[\eg][]{Hopkins06a, Madau14a}.
A leading explanation for this decline is the decrease in molecular gas content and
\SF efficiency \citep[]{Erb06a, CW13, Walter14a},
but confirmation from direct molecular gas measurements at intermediate redshift
$(0.2\,<\,z\,<\,1)$ is still limited both in number and in the variety of galaxy populations probed ---
only unresolved CO observations have been carried out for $\sim$$30$ ultra-luminous infrared galaxies (ULIRGs) up to date \citep{Combes11a, Combes13a}. %18+15

Meanwhile, empirical scaling relations such as the
$M_{\rm BH}$$-$\mbulge relation \citep[e.g.,][]{Magorrian98a, HR04a}
has been established locally, suggesting a co-eval growth between local SMBHs and their host galaxies.
Attempts to extend this relation out to higher redshifts, beyond the peak epoch
of \SF and AGN activity, have been made in recent years.
These studies find that 
% the $M_{\rm BH}$$/$$M_{\rm bulge}$ ratios of high-$z$ submillimeter galaxies (SMGs) 
% are lower than those of local AGN host galaxies \citep[\eg][]{Borys05a,Alexander08a}, 
% whereas those of high-$z$ quasar host galaxies 
% are higher than the local values \citep[\eg][]{Walter04a, McLure06a,Peng06a, Riechers08a,Coppin08a}, 
% suggesting that the co-eval growth of high-$z$ AGNs and their host galaxies follows different 
% timescales compared to their local counterparts. % different AGN mechanism, different AGN accretion rate
% However, any links between such fast-growing SMBHs and their host galaxies are not yet established at these high redshifts.
high-$z$ AGN host galaxies do not follow the same $M_{\rm BH}$$-$\mbulge relation \citep[\eg][]{Walter04a,Borys05a, McLure06a,Peng06a, Riechers08a,Coppin08a,Alexander08a}.
Yet, the $M_{\rm BH}$$-$\mbulge relation remains poorly-constrained 
at intermediate redshifts.
This is owed to the difficulties in measuring stellar emission in optical quasar host galaxies, where
emission of quasars produces sidelobes that saturate the lower surface brightness starlight in the host galaxy,
or alternatively, due to the lack of spatially resolved CO observations to provide dynamical constraints on the stellar 
(bulge) mass.
With strong gravitational lensing, emission from the background quasar and its host galaxy originating from 
different source-plane locations and of different extent is being magnified and stretched out by different amount.
The lensed emission is therefore spatially resolved over more beams which reduces the sidelobes and enables 
measurements of the stellar emission in the host galaxy.

\defcitealias{Sluse03a}{S03}
The quasar RXS J113151.62$-$123158 (hereafter RXJ1131)
at $z_\textrm{s, QSO}$\eq$0.685$ \citep[hereafter S03]{Sluse03a} is a particularly well-suited source for
studying the evolution of molecular gas properties in quasar host galaxies and the
connection between SMBHs and their host galaxies at intermediate redshift
given its unique lensing configuration.
The stellar emission in the host galaxy of RXJ1131 is lensed into
an Einstein ring of $1$\farcs$83$ in radius
that is clearly separated from the quadruply imaged quasar emission \citep[hereafter C06]{Claeskens06a}.
The foreground lens is an elliptical galaxy at $z_\textrm{L}$\eq$0.295$ \citepalias{Sluse03a}.
 \citet[][]{Reis14a} report a high spin parameter of $a$\,$\sim$\,$0.9$ for the moderate-mass black hole in RXJ1131 \citep[$M_{\rm BH}$\eq$8$\E{7}\Msun;][]{Sluse12a}, %based on H beta line
with an intrinsic bolometric luminosity of $L_{\rm bol, X}$\eq$1.3$\E{45}\, ergs\,s\pmOne \citep{Pooley07a}.

In this paper, we present \bco and \cco line data observed towards RXJ1131 and
broadband photometry spanning rest-frame UV to radio wavelengths to 
study the properties of the molecular gas, the dust and stellar populations in the 
host galaxy of RXJ1131.
In \Sec{obs}, we outline details of the observations and of our data reduction procedures.
In \Sec{HST}, we describe the astrometric corrections
applied to an image taken with the \hst ({\it HST}) by aligning it to
a radio image taken with the \vla (VLA).
In \Sec{results}, we report measurements of the \bco and \cco lines and broadband
photometry in RXJ1131.
In \Sec{anal}, we present dynamical lens modeling of the \bco data and
spectral energy distribution (SED) modeling of the photometric data.
In \Sec{diss}, we discuss 
the ISM properties of the host galaxy of RXJ1131 and compare 
them to other galaxy populations at low and high redshift.
Finally, we summarize the main results and present our conclusions in \Sec{sum}.
We use a concordance $\Lambda$CDM cosmology throughout this paper, with
parameters from the WMAP9 results:
$H_0$ = $69.32$ \kms Mpc\pmOne, $\Omega_{\rm M}$ = $0.29$, and
$\Omega_{\Lambda}$ = $0.71$ \citep{Hinshaw13a}.
% same as LR16a: i.e. use last column (WMAP+eCMB+BAO+H0) of Tab 4 in Hinshaw+13 because that's what astropy use for WMAP9 as cosmo
\defcitealias{Claeskens06a}{C06}

%--------------------------------------------------------------------------
%                          Observations details
%--------------------------------------------------------------------------
\section{Observations} \label{sec:obs}
\subsection{PdBI \bco}
Observations of the \bco rotational line
($\nu_{\rm rest}$\,=\,$230.53800$ GHz) 
at the $\nu_{\rm LO}$\eq139.256\,GHz
% redshifted to $\nu_{\rm obs}$\eq139.40739\,GHz
were carried out using IRAM \pdbi (PdBI; Program ID: S14BX; PI: D.
Riechers). Two observing runs were carried out on 2014 December 06 and 2015
February 05 under good weather conditions in the C and D array configurations,
respectively. 
This resulted in 3.75 hours of cumulative six antenna-equivalent on-source
 time after discarding unusable visibility data.
The 2 mm receivers were used to cover the redshifted \bco line
and the underlying continuum emission, employing a correlator setup that provides
an effective bandwidth of 3.6 GHz (dual polarization) and a spectral resolution of 10.0 MHz ($\sim$
$21.5$ \kms).  {\bf DR: DSB, 320 MHz/128 channels? <-> 2.5 MHz? I did not bin the data}
The nearby quasars B1127$-$145 and B1124$-$186 were observed every 22 minutes
for pointing, secondary amplitude, and phase calibration, and B1055$+$018 was
observed as the bandpass calibrator for both tracks.
MWC349 and 3C279 were observed as primary flux calibrators for the C and D
array observations, respectively, yielding calibration accuracy better than 15\%.

@@

\subsection{CARMA \cco}
Observations of the \cco rotational line in RXJ1131
($\nu_{\rm rest}$\,=\,345.79599\,GHz) redshifted to $\nu_{\rm obs}$\eq209.10443\,GHz
were carried out with the \carma (CARMA;
Program ID: cf0098; PI: D. Riechers)
in the D array configuration on 2014 February 02 under poor 1.5\,mm
weather conditions and on 2014 February 17 under good 1.5\,mm
weather conditions. 
This resulted in a total on-source time of 2.94 hours after flagging poor
visibility data.
The correlator setup provides a bandwidth of 3.75 GHz in
each sideband and a %native
spectral resolution of 12.5 MHz ($\sim$17.9 \kms). The
line was placed in the lower sideband with the local oscillator tuned to $\nu_{\rm LO}\sim$216 GHz. The radio quasars J1127$-$189 (first track) and 3C273
(second track) were observed
every 15 minutes for pointing, amplitude, and phase calibration. Mars was
observed as the primary absolute flux calibrator and 3C279 was observed as
the bandpass calibrator for both tracks.

% poor phase note
% /Users/admin/Research/RXJ1131/CARMA/imagingD23/checkQuality.csh
% rms scatter over ~ 90 klambda
% max baseline of obs. is 145 m
%
% calflux source=1127-189 in=$MIRCAT/FluxSource.cat device=/xs
% Flux of: 1127-189  14FEB04.50 at 227.0 GHz:  0.65 Jy; rms: 0.10 Jy
% Extrapolate to 215.673 GHz with spectral index = -0.986 --> 0.684 Jy
% MIRIAD from bootstrap: Median Flux @ 215.673 GHz:     0.710 --> 3%
%
% amp de-correlation: % loss scales = 1- exp(-(50º*pi/180)^2 / 2) = ~32% loss
Given that the phase calibrator used for the first track was faint and was
observed under poor weather conditions and that the phase calibrator used for
the second track was far from our target source, the phase calibration is
subpar, with an rms scatter $\sim$50$\degr$ over a baseline length of $\sim$135\,m.
We thus conservatively estimate
a calibration accuracy of $\sim$35\% based on the flux scale uncertainties,
the gain variations over time, and the phase scatter on the calibrated data. We
therefore treat the derived \cco line intensity with caution and ensure that our physical interpretation
of this system and the conclusion of this paper do not rely on this quantity.

@@

\subsection{VLA (Archival)} 
Our analysis also uses archival data of the 4.885\,GHz %4.88510 GHz
radio continuum obtained with the
VLA (Program ID: AW741; PI: Wucknitz).
Observations were carried out on 2008 December 29 under excellent weather
conditions in the A array configuration for a total of $\sim$7 hours on-source time. The C-band receivers were used with a continuum mode setup,
providing a bandwidth of 50 MHz for the two IF bands with full polarization.
The nearby radio quasar J1130$-$149 was observed every 10 minutes for
pointing, amplitude, and phase calibration. J1331$+$305 was observed as the
primary flux calibrator, and J0319$+$415 was observed as the bandpass
calibrator, yielding $\sim$10\% calibration accuracy.
We used \ncode{aips} to calibrate the visibility data.
The calibrated visibility data were imaged and deconvolved using
the CLEAN algorithm with robust\,=\,0, which
was chosen to obtain a higher resolution image given high SNR.
This yields a synthesized clean
beam size of 0$\farcs$49 $\times$ 0\farcs35 (PA\,=\,0.18$\degr$) and a final
rms noise of $\sigma$ = 13 $\mu$Jy beam\pmOne.

\section{HST astrometry} \label{sec:HST}
We obtained an {\it HST} image taken with
the ACS
using the F555W filter ($V$-band)
from the
Hubble Legacy Archive.
The details of the observations can be found
in \citetalias{Claeskens06a}.
We adopt the VLA 5\,GHz map as the
reference coordinate frame to align the optical ($V$-band) image.
We shift the latter to the east by 0\farcs5963 in R.A. and $+$0\farcs8372 in
Dec., which is consistent with the typical astrometric precision (1$^{\prime\prime}-$2$^{\prime\prime}$) of
images from the Hubble Legacy Archive\footnote{http://hla.stsci.edu/hla\_faq.
html}. This astrometric correction is critical to avoid artificial spatial
offsets between different emitting regions and to carry out our lens modeling,
in which the absolute position of the foreground lensing galaxy is based on
its coordinates in the high-resolution optical image.
The VLA image is calibrated using a well-monitored phase
calibrator, with absolute positional accuracy of $\sim$2 mas.
For this reason, the absolute alignment between the VLA image and other
interferometric images reported in this paper are expected to have an astrometric
precision better than 0\farcs1, modulo uncertainties related to the SNR and phase
instability.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.45\textwidth]{../Figures/SpecCO21_twinx.eps}
\caption{Continuum-subtracted spectrum (histogram) of \bco emission toward RXJ1131, with a spectral resolution of 
21.5\,\kms. The solid black line shows a double-Gaussian fit to the line profile.
The velocity scale is with respect to $z$\,=\,0.6537, which is approximately the line center
considering the asymmetry resulting from differential lensing.
A detailed discussion of this effect is presented in
\Sec{differential}.
 \label{fig:CO21spec}}
\end{figure}

% moment 0 map, highest SNR: CASA chan=125~159 <=> GILDAS 126-160 <=> python 125-160
% - sigma = 0.305 Jy km/s/Beam
%   - note that theoretical sigma is lower = 1.5 * sqrt(160-126+1) * 21.5 ~ 0.2 Jy km/s /beam, due to higher noise in some channels with emission
We construct a zeroth order moment map, red/blue channel maps, and
first and second moment maps, as shown in \Fig{CO21mom},
using the $uv$-continuum subtracted data cube over a velocity range of
$\Delta v$ $\sim$ 750\,\kms.
The higher-order moment maps are produced using
unbinned channel maps with 3$\sigma$ clipping.
The peak flux density is 8.12\pmm0.30 Jy\,\kms\,beam\pmOne
in the intensity map.

The deconvolved source size FWHM obtained by fitting a two-dimensional Gaussian to the integrated line map in
the image plane
is 5\farcs1\pmm0\farcs72 $\times$ 3\farcs72\pmm0\farcs66,
and thus, the emission is resolved over $\sim$2.2 beams.
While the lensed emission is not strictly distributed as a two-dimensional
Gaussian;
the fit recovers the line intensity enclosed by the emitting
region, therefore we take this as an estimate on the full extent of the lensed
emission. On the other hand, if we assume that the spatial distribution of
the lensed molecular gas emission is similar to that in the optical to \nir
wavelengths, the lensed emission would be more accurately described by an
annulus, enclosing the partially complete ``Einstein ring'' and
the lensed knots (see \Fig{CO21mom}).
